---
title: "Predicting Grain Crop Yields in India for Years 1997 - 2020"
output: html_document
date: "2023-07-31"
author: "Team BCG: Cameron Greenwalt, Mark Bauer, Ben Carman"
---

```{r}
# TODO: Remove TODO's
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libariries and Functions

Below are some libraries and helper functions we utilize

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(plyr)
library(glmnet)
library(MASS)
library(MASSExtra)
library(car)
library(lmtest)
library(coefplot)
library(rlist)
```

```{r}
plot_fitted_resid <- function(model, pointcol = "dodgerblue", linecol = "darkorange") {
  plot(fitted(model), resid(model),
    col = pointcol, pch = 20, cex = 1.5,
    xlab = "Fitted", ylab = "Residuals", main = "Fitted vs. Residuals"
  )
  abline(h = 0, col = linecol, lwd = 2)
}

plot_qq <- function(model, pointcol = "dodgerblue", linecol = "darkorange") {
  qqnorm(resid(model), col = pointcol, pch = 20, cex = 1.5)
  qqline(resid(model), col = linecol, lwd = 2)
}

plot_fitted_v_actual <- function(model, pointcol = "dodgerblue", linecol = "darkorange") {
  set_plot_grid(1, 1)
  plot(fitted(model), resid(model) + fitted(model),
    col = pointcol, pch = 20, cex = 1.5,
    xlab = "Fitted", ylab = "Actual", main = "Fitted vs. Actual"
  )
  lines(x = fitted(model), y = fitted(model), col = linecol, lwd = 2)
}

set_plot_grid <- function(nrow = 1, ncol = 1) {
  par(mfrow = c(nrow, ncol))
}

assumptions_plots <- function(model) {
  set_plot_grid(1, 3)
  plot_fitted_resid(model)
  plot_qq(model)
  hist(resid(model), main = "Histogram of Model Residuals", xlab = "Residuals")
}

plot_cooksd <- function(model) {
  set_plot_grid(1, 1)
  cooksd <- cooks.distance(model)
  n <- length(model$residuals)
  infl <- cooksd > (4 / n)
  plot(1:n,
    cooksd,
    col = ifelse(infl == 1, "red", "black"),
    xlab = "Sample Index",
    ylab = "Cook's Distance",
    main = "Cook's Distances"
  )
  abline(
    h = 4 / n,
    col = "royalblue"
  )
}

bc_transform <- function(x, lambda) {
  #((x)^lambda - 1) / lambda
  MASSExtra::bc(x, lambda)
}

get_loocv_rmse <- function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model)))^2))
}

get_adj_r2 <- function(model) {
  summary(model)$adj.r.squared
}

idx_infl_pts <- function(model) {
  cooks.distance(model) > (4 / length(resid(model)))
}

output_diagnostics = function(model, vif = TRUE) {
  if (length(coef(model)) < 30) {
    print("Model Summary:")
    print(summary(model))
  }
  plot_fitted_v_actual(model)
  assumptions_plots(model)
  plot_cooksd(model)
  bptest(model)
  shapiro.test(resid(model))
  
  print(c("LOOCV-RMSE:", get_loocv_rmse(model)))
  
  if (vif) {
    car::vif(model)
  }
}

inverse_bc_transform = function(x, lambda) {
  # (x*λ+1)^(1/λ)
  #((x*lambda)+1)^(1/lambda)
  bc_inv(x, lambda)
}
```

---

## Introduction

In this project we will explore our ability to predict the amount of production of grain crops in India based on some information about the land in which they were grown on (field size, general region), the type of grain, inputs used (fertilizer, pesticide), and weather (season, rain).

Below we provide a summary or the data.  The source for the data can be found (here)[https://www.kaggle.com/datasets/akshatgupta7/crop-yield-in-indian-states-dataset].  

Response - Production: The quantity of crop production (in metric tons).

Predictors

- Crop: The name of the cultivated crop .  
- Crop_Year: The year in which the crop was grown.
- Season: The specific cropping season (e.g., Kharif, Rabi, Whole Year).
- State: The Indian state where the crop was cultivated.
- Area: The total land area (in hectares) under cultivation for the specific crop.
- Annual_Rainfall: The annual rainfall received in the crop-growing region (in mm).
- Fertilizer: The total amount of fertilizer used for the crop (in kilograms).
- Pesticide: The total amount of pesticide used for the crop (in kilograms).

## Methods

### Data import, exploration, and cleaning

```{r message=FALSE}
crops <- read_csv("crop_yield.csv")
```

It turns out (as mentioned above) our original intended response, `Yield`, is a metric derived by computing `Production / Area`. It would be incorrect for us to include `Production` and `Area` as predictors in a model whose response is `Yield`. Instead, we remove `Yield` and use the remaining variables to predict `Production`.

```{r}
crops <- crops[, !(names(crops) %in% c("Yield"))]
```

In order to limit scope, we also limit the types of crops under observation to a subset of grains (e.g., wheat, barley, rye, etc.). We also convert categorical variables to factor variables as appropriate.

```{r message=FALSE, warning=FALSE}
grains <- c(
  "Amaranth", "Barley", "Bulgur", "Corn", "Farro", "Einkorn", "Emmer",
  "Spelt", "Millet", "Freaked", "Farik", "Durum Wheat",
  "Khorasan Wheat", "Oats", "Quinoa", "Kana", "Rice", "Rye",
  "Sorghum", "Teff", "Triticale", "Wheat", "Wild Rice", "Wheat",
  "Wheat berries", "Hominy", "Spelt", "Rye", "Brown Rice", "Farro",
  "Emmer", "Barley", "Bran", "Durum Wheat ", "Triticale",
  "Bulgur Wheat", "Couscous", "Farina", "Kamut", "Orzo", "Semolina",
  "Graham", "Oats", "Corn", "maize", "Cornflour ", "Cornmeal", "Rice",
  "Wild Rice", "Teff", "Montina flour", "Sorghum", "Oats", "Freekeh",
  "Emmer", "Eikorn", "Malt", "Graham", "Couscous", "Polenta", "Muesli",
  "Seitan", "Panko", "Grain Alcohol", "Atta Flour", "Amaranth", "Quinoa",
  "Millet", "Barley Malt", "Beer", "Bleached Flour", "Breads",
  "Baked Goods", "Brown Rice Syrup", "Buckwheat", "Corn Flakes",
  "Croutons", "Cereals", "Wheat Germ", "Enriched Bleach Flour",
  "Malted Barley Flour", "Millet", "Granary Flour", "Groats",
  "Pastas", "Matzo", "Rice Milk", "Seitan", "Tabbouleh", "Udon",
  "Corn Starch", "Wheat nuts", "Other Cereals"
)

# Drop non-grain rows
crops <- crops[str_squish(tolower(crops$Crop)) %in% str_squish(tolower(grains)), ]

# Convert categorical variables to factor variables
crops$Crop <- as.factor(crops$Crop)
crops$Season <- as.factor(crops$Season)
crops$State <- as.factor(crops$State)

levels(crops$Crop)
```

Here we present a few rows from the cleaned and pared down dataset:

```{r}
kable(
  head(crops, 5)
)
```

As we prepare the dataset for modeling, we also explore the remaining factor variables:

```{r}
levels(crops$Season)
```

The season column appears reasonable. Let's also check the values for State:

```{r}
levels(crops$State)
```

In order to keep our model's complexity minimal, we will pre-process the data to consider zones in India instead of working at the granularity of individual states:

```{r}
# See https://en.wikipedia.org/wiki/Administrative_divisions_of_India

states_to_zones <- c(
  "Chandigarh" = "Northern",
  "Delhi" = "Northern",
  "Haryana" = "Northern",
  "Himachal Pradesh" = "Northern",
  "Jammu and Kashmir" = "Northern",
  "Ladakh" = "Northern",
  "Punjab" = "Northern",
  "Rajasthan" = "Northern",
  "Assam" = "North Eastern",
  "Arunachal Pradesh" = "North Eastern",
  "Manipur" = "North Eastern",
  "Meghalaya" = "North Eastern",
  "Mizoram" = "North Eastern",
  "Nagaland" = "North Eastern",
  "Tripura" = "North Eastern",
  "Sikkim" = "North Eastern",
  "Chhattisgarh" = "Central",
  "Madhya Pradesh" = "Central",
  "Uttarakhand" = "Central",
  "Uttar Pradesh" = "Central",
  "Bihar" = "Eastern",
  "Jharkhand" = "Eastern",
  "Odisha" = "Eastern",
  "West Bengal" = "Eastern",
  "Goa" = "Western",
  "Gujarat" = "Western",
  "Maharashtra" = "Western",
  "Andhra Pradesh" = "Southern",
  "Karnataka" = "Southern",
  "Kerala" = "Southern",
  "Puducherry" = "Southern",
  "Tamil Nadu" = "Southern",
  "Telangana" = "Southern"
)


crops$State <- revalue(crops$State, states_to_zones)
names(crops)[which(names(crops) == "State")] <- "Zone"
crops$Zone <- as.factor(crops$Zone)
levels(crops$Zone)
```

```{r}
# Train-test split
set.seed(0)

sample <- sample(c(TRUE, FALSE), nrow(crops), replace = TRUE, prob = c(0.8, 0.2))
crops.train <- crops[sample, ]
crops.test.cropyear_cont <- crops[!sample, ]

nrow(crops)
nrow(crops.train)
nrow(crops.test.cropyear_cont)
```

As we examined the data and discussed, something interesting we wanted to test/investigate is the concept of crop year as a categorical versus a continuous variable.  For our initial try, we utilize it as continuous.  

```{r fig.height=5, fig.width=10}
vars.res <- "Production"
vars.con <- c("Area", "Annual_Rainfall", "Fertilizer", "Pesticide", "Crop_Year")
vars.cat <- c("Crop", "Season", "Zone")
```

We then do a quick examination of the pairs plot.  

```{r}
pairs(crops.train[, c(vars.con, vars.res)])
```

```{r message=FALSE}
# TODO: Add some inferences from examining the pairs plot?
# colinearity with Fert and Pest?
```

For our first model, we chose a baseline additive model with all predictors.

```{r fig.height=5, fig.width=8}
baseline.cropyear_cont <- lm(Production ~ ., data = crops.train)
output_diagnostics(baseline.cropyear_cont)
```

We have one influential point we chose to remove.  While we are not completely confident in our justification for this, we are prioritizing simplicity over rigor in both this removal, and in other instances of high influential points that are found.  

```{r fig.height=5, fig.width=8}
idx_remove <- which.max(cooks.distance(baseline.cropyear_cont))
crops.train.no_infl <- crops.train[-idx_remove, ]

baseline.no_infl.cropyear_cont <- lm(Production ~ ., data = crops.train.no_infl)
output_diagnostics(baseline.no_infl.cropyear_cont)
```

Our diagnostics gave us reason for concern with regard to our LINE assumptions, so we perform a Box-Cox transformation.  

```{r fig.height=5, fig.width=8}
# See https://www.statology.org/box-cox-transformation-in-r/

# Define plotting area
set_plot_grid(1, 2)

# Box-cox requires that the response be positive (0 doesn't work), so we add
# 1 here (just need to remember to subtract 1 when giving actual predictions)
crops$Production = crops$Production + 1
crops.train$Production = crops.train$Production + 1
crops.test.cropyear_cont$Production = crops.test.cropyear_cont$Production + 1
crops.train.no_infl$Production = crops.train.no_infl$Production + 1

baseline.cropyear_cont <- lm(I(Production) ~ ., data = crops.train.no_infl)

bc <- boxcox(baseline.cropyear_cont)
(baseline.bc.cropyear_cont_lambda <- bc$x[which.max(bc$y)])

baseline.bc.cropyear_cont <- lm(bc_transform(Production, baseline.bc.cropyear_cont_lambda) ~ ., data = crops.train.no_infl)
output_diagnostics(baseline.bc.cropyear_cont)
```

***

Next, we move from our base additive model to one with interactions between the categorical variables.  We use a similar approach to examine and evaluate the model.  

```{r}
# continuous-categorical interactions
baseline.poly.form.cropyear_cont = as.formula(Production ~ 
                        (poly(Area, degree = 2, raw = 1) +
                        poly(Annual_Rainfall, degree = 2, raw = 1) +
                        poly(Fertilizer, degree = 2, raw = 1) +
                        poly(Pesticide, degree = 2, raw = 1) +
                        poly(Crop_Year, degree = 2, raw = 1)) *
                        (Crop + Season + Zone))
baseline.poly.cropyear_cont = lm(baseline.poly.form.cropyear_cont, data = crops.train)
output_diagnostics(baseline.poly.cropyear_cont, vif = 0)
```

```{r}
crops.train.poly.no_infl <- crops.train[!idx_infl_pts(baseline.poly.cropyear_cont),]

baseline.poly.cropyear_cont = lm(baseline.poly.form.cropyear_cont, data = crops.train.poly.no_infl)
output_diagnostics(baseline.poly.cropyear_cont, vif = 0)
```

```{r}
bc <- boxcox(baseline.poly.cropyear_cont)
(baseline.poly.bc.cropyear_cont_lambda <- bc$x[which.max(bc$y)])

baseline.poly.bc.cropyear_cont = lm(paste("bc_transform(Production, baseline.poly.bc.cropyear_cont_lambda) ~ ", 
                             paste(attr(terms(baseline.poly.form.cropyear_cont), 
                                        which = "term.labels"), 
                                   collapse = " + ")),
                        data = crops.train.poly.no_infl)
output_diagnostics(baseline.poly.bc.cropyear_cont, vif = 0)
```

***

Continuing on, we evaluate a model with other transformations of the predictors, including degree two, square roots, and logs of the continuous predictors.  

```{r}
multi.form.cropyear_cont = as.formula(Production ~ 
                        (poly(Area, degree = 2, raw = 1) +
                        poly(Annual_Rainfall, degree = 2, raw = 1) +
                        poly(Fertilizer, degree = 2, raw = 1) +
                        poly(Pesticide, degree = 2, raw = 1) +
                        poly(Crop_Year, degree = 2, raw = 1) +
                        I(sqrt(Area)) +
                        I(sqrt(Annual_Rainfall)) +
                        I(sqrt(Fertilizer)) +
                        I(sqrt(Pesticide)) +
                        I(sqrt(Crop_Year)) +
                        I(log(Area)) +
                        I(log(Annual_Rainfall)) +
                        I(log(Fertilizer)) +
                        I(log(Pesticide)) +
                        I(log(Crop_Year))) *
                        (Crop + Season + Zone))
multi.full.cropyear_cont = lm(multi.form.cropyear_cont, data = crops.train)
output_diagnostics(multi.full.cropyear_cont, vif = FALSE)
```

```{r}
crops.train.multi.no_infl <- crops.train[!idx_infl_pts(multi.full.cropyear_cont),]

multi.full.cropyear_cont = lm(multi.form.cropyear_cont, data = crops.train.multi.no_infl)
output_diagnostics(multi.full.cropyear_cont, vif = FALSE)
```

```{r}
bc <- boxcox(multi.full.cropyear_cont)
(multi.full.bc.cropyear_cont_lambda <- bc$x[which.max(bc$y)])

multi.full.bc.cropyear_cont = lm(paste("bc_transform(Production, multi.full.bc.cropyear_cont_lambda) ~ ",
                             paste(attr(terms(multi.form.cropyear_cont),
                                        which = "term.labels"),
                                   collapse = " + ")),
                        data = crops.train.multi.no_infl)
output_diagnostics(multi.full.bc.cropyear_cont, vif = FALSE)
```

***

We also want to evaluate whether our models can be made simpler, so we do variable selection using BIC on them

```{r echo=FALSE, message=FALSE}
# multi.bic.backward.bc.cropyear_cont <- step(multi.full.bc.cropyear_cont,
#   direction = "backward", trace = 0,
#   k = log(nrow(crops.train.multi.no_infl))
# )
# output_diagnostics(multi.bic.backward.bc.cropyear_cont, vif = 0)
```

```{r}
multi.bic.forward.bc.cropyear_cont <- step(lm(bc_transform(Production, multi.full.bc.cropyear_cont_lambda) ~ 1, data = crops.train.multi.no_infl),
                                           scope = as.formula(
                                             paste("bc_transform(Production, multi.full.bc.cropyear_cont_lambda) ~ ",
                             paste(attr(terms(multi.form.cropyear_cont),
                                        which = "term.labels"),
                                   collapse = " + "))
                                           ),
                             direction = "forward", trace = 0,
  k = log(nrow(crops.train.multi.no_infl))
)
output_diagnostics(multi.bic.forward.bc.cropyear_cont, vif = 0)
```

***

As mentioned above, one point of interest as we explored our models was whether crop year was better as a continuous or categorical variable.  Below we perform a set of similar tests, only with crop year as a category.  

```{r fig.height=5, fig.width=10}
vars.res <- "Production"
vars.con <- c("Area", "Annual_Rainfall", "Fertilizer", "Pesticide")
vars.cat <- c("Crop", "Season", "Zone", "Crop_Year")
```

```{r}
crops$Crop_Year = as.factor(crops$Crop_Year)
```


```{r}
# Train-test split
set.seed(0)

sample <- sample(c(TRUE, FALSE), nrow(crops), replace = TRUE, prob = c(0.8, 0.2))
crops.train <- crops[sample, ]
crops.test.cropyear_cat <- crops[!sample, ]

nrow(crops)
nrow(crops.train)
nrow(crops.test.cropyear_cat)
```

```{r}
baseline.cropyear_cat <- lm(Production ~ ., data = crops.train)
output_diagnostics(baseline.cropyear_cat)
```

```{r}
idx_remove <- which.max(cooks.distance(baseline.cropyear_cat))
crops.train.no_infl <- crops.train[-idx_remove, ]

baseline.no_infl.cropyear_cat <- lm(Production ~ ., data = crops.train.no_infl)
output_diagnostics(baseline.no_infl.cropyear_cat)
```


```{r}
baseline.cropyear_cat <- lm(I(Production) ~ ., data = crops.train.no_infl)

bc <- boxcox(baseline.cropyear_cat)
(baseline.bc.cropyear_cat_lambda <- bc$x[which.max(bc$y)])

baseline.bc.cropyear_cat <- lm(bc_transform(Production, baseline.bc.cropyear_cat_lambda) ~ ., data = crops.train.no_infl)
output_diagnostics(baseline.bc.cropyear_cat)
```


```{r}
# continuous-categorical interactions
baseline.poly.form.cropyear_cat = as.formula(Production ~ 
                        (poly(Area, degree = 2, raw = 1) +
                        poly(Annual_Rainfall, degree = 2, raw = 1) +
                        poly(Fertilizer, degree = 2, raw = 1) +
                        poly(Pesticide, degree = 2, raw = 1)) *
                        (Crop + Season + Zone + Crop_Year))
baseline.poly.cropyear_cat = lm(baseline.poly.form.cropyear_cat, data = crops.train)
output_diagnostics(baseline.poly.cropyear_cat, vif = 0)
```

```{r}
crops.train.poly.no_infl <- crops.train[!idx_infl_pts(baseline.poly.cropyear_cat),]

baseline.poly.cropyear_cat = lm(baseline.poly.form.cropyear_cat, data = crops.train.poly.no_infl)
output_diagnostics(baseline.poly.cropyear_cat, vif = 0)
```

```{r}
bc <- boxcox(baseline.poly.cropyear_cat)
(baseline.poly.bc.cropyear_cat_lambda <- bc$x[which.max(bc$y)])

baseline.poly.bc.cropyear_cat = lm(paste("bc_transform(Production, baseline.poly.bc.cropyear_cat_lambda) ~ ", 
                             paste(attr(terms(baseline.poly.form.cropyear_cat), 
                                        which = "term.labels"), 
                                   collapse = " + ")),
                        data = crops.train.poly.no_infl)
output_diagnostics(baseline.poly.bc.cropyear_cat, vif = 0)
```

```{r}
multi.form.cropyear_cat = as.formula(Production ~ 
                        (poly(Area, degree = 2, raw = 1) +
                        poly(Annual_Rainfall, degree = 2, raw = 1) +
                        poly(Fertilizer, degree = 2, raw = 1) +
                        poly(Pesticide, degree = 2, raw = 1) +
                        I(sqrt(Area)) +
                        I(sqrt(Annual_Rainfall)) +
                        I(sqrt(Fertilizer)) +
                        I(sqrt(Pesticide)) +
                        I(log(Area)) +
                        I(log(Annual_Rainfall)) +
                        I(log(Fertilizer)) +
                        I(log(Pesticide))) *
                        (Crop + Season + Zone + Crop_Year))
multi.full.cropyear_cat = lm(multi.form.cropyear_cat, data = crops.train)
output_diagnostics(multi.full.cropyear_cat, vif = FALSE)
```

```{r}
crops.train.multi.no_infl <- crops.train[!idx_infl_pts(multi.full.cropyear_cat),]

multi.full.cropyear_cat = lm(multi.form.cropyear_cat, data = crops.train.multi.no_infl)
output_diagnostics(multi.full.cropyear_cat, vif = FALSE)
```

```{r}
bc <- boxcox(multi.full.cropyear_cat)
(multi.full.bc.cropyear_cat_lambda <- bc$x[which.max(bc$y)])

multi.full.bc.cropyear_cat = lm(paste("bc_transform(Production, multi.full.bc.cropyear_cat_lambda) ~ ",
                             paste(attr(terms(multi.form.cropyear_cat),
                                        which = "term.labels"),
                                   collapse = " + ")),
                        data = crops.train.multi.no_infl)
output_diagnostics(multi.full.bc.cropyear_cat, vif = FALSE)
```

```{r echo=FALSE, message=FALSE}
# multi.aic.backward.bc.cropyear_cat <- step(multi.full.bc.cropyear_cat, direction = "backward", trace = 0)
# output_diagnostics(multi.aic.backward.bc.cropyear_cat, vif = 0)
```

```{r echo=FALSE, message=FALSE}
# multi.bic.backward.bc.cropyear_cat <- step(multi.full.bc.cropyear_cat,
#   direction = "backward", trace = 0,
#   k = log(nrow(crops.train.multi.no_infl))
# )
# output_diagnostics(multi.bic.backward.bc.cropyear_cat, vif = 0)
```

```{r echo=FALSE, message=FALSE}
# multi.aic.forward.bc.cropyear_cat <- step(lm(bc_transform(Production, multi.full.bc.cropyear_cat_lambda) ~ 1, data = crops.train.multi.no_infl),
#                                            scope = as.formula(
#                                              paste("bc_transform(Production, multi.full.bc.cropyear_cat_lambda) ~ ",
#                              paste(attr(terms(multi.form.cropyear_cat),
#                                         which = "term.labels"),
#                                    collapse = " + "))
#                                            ),
#                              direction = "forward", trace = 0)
# output_diagnostics(multi.aic.forward.bc.cropyear_cat, vif = 0)
```

```{r}
multi.bic.forward.bc.cropyear_cat <- step(lm(bc_transform(Production, multi.full.bc.cropyear_cat_lambda) ~ 1, data = crops.train.multi.no_infl),
                                           scope = as.formula(
                                             paste("bc_transform(Production, multi.full.bc.cropyear_cat_lambda) ~ ",
                             paste(attr(terms(multi.form.cropyear_cat),
                                        which = "term.labels"),
                                   collapse = " + "))
                                           ),
                             direction = "forward", trace = 0,
  k = log(nrow(crops.train.multi.no_infl))
)
output_diagnostics(multi.bic.forward.bc.cropyear_cat, vif = 0)
```

The next iteration of our crop year experiment was to leave the potential predictor out entirely.  We then proceed to run the similar set of evaluations one more time.  

```{r fig.height=5, fig.width=10}
vars.res <- "Production"
vars.con <- c("Area", "Annual_Rainfall", "Fertilizer", "Pesticide")
vars.cat <- c("Crop", "Season", "Zone")
```


```{r}
# Train-test split
set.seed(0)

sample <- sample(c(TRUE, FALSE), nrow(crops), replace = TRUE, prob = c(0.8, 0.2))
crops.train <- crops[sample, ]
crops.test.cropyear_omit <- crops[!sample, ]

crops.train <- crops.train[, !(names(crops.train) %in% c("Crop_Year"))]
crops.test.cropyear_omit <- crops.test.cropyear_omit[, !(names(crops.test.cropyear_omit) %in% c("Crop_Year"))]

nrow(crops)
nrow(crops.train)
nrow(crops.test.cropyear_omit)
```

```{r fig.height=5, fig.width=8}
baseline.cropyear_omit <- lm(Production ~ ., data = crops.train)
output_diagnostics(baseline.cropyear_omit)
```

```{r fig.height=5, fig.width=8}
idx_remove <- which.max(cooks.distance(baseline.cropyear_omit))
crops.train.no_infl <- crops.train[-idx_remove, ]

baseline.no_infl.cropyear_omit <- lm(Production ~ ., data = crops.train.no_infl)
output_diagnostics(baseline.no_infl.cropyear_omit)
```


```{r}
baseline.cropyear_omit <- lm(I(Production) ~ ., data = crops.train.no_infl)

bc <- boxcox(baseline.cropyear_omit)
(baseline.bc.cropyear_omit_lambda <- bc$x[which.max(bc$y)])

baseline.bc.cropyear_omit <- lm(bc_transform(Production, baseline.bc.cropyear_omit_lambda) ~ ., data = crops.train.no_infl)
output_diagnostics(baseline.bc.cropyear_omit)
```

```{r}
# continuous-categorical interactions
baseline.poly.form.cropyear_omit = as.formula(Production ~ 
                        (poly(Area, degree = 2, raw = 1) +
                        poly(Annual_Rainfall, degree = 2, raw = 1) +
                        poly(Fertilizer, degree = 2, raw = 1) +
                        poly(Pesticide, degree = 2, raw = 1)) *
                        (Crop + Season + Zone))
baseline.poly.cropyear_omit = lm(baseline.poly.form.cropyear_omit, data = crops.train)
output_diagnostics(baseline.poly.cropyear_omit, vif = 0)
```

```{r}
crops.train.poly.no_infl <- crops.train[!idx_infl_pts(baseline.poly.cropyear_omit),]

baseline.poly.cropyear_omit = lm(baseline.poly.form.cropyear_omit, data = crops.train.poly.no_infl)
output_diagnostics(baseline.poly.cropyear_omit, vif = 0)
```

```{r}
bc <- boxcox(baseline.poly.cropyear_omit)
(baseline.poly.bc.cropyear_omit_lambda <- bc$x[which.max(bc$y)])

baseline.poly.bc.cropyear_omit = lm(paste("bc_transform(Production, baseline.poly.bc.cropyear_omit_lambda) ~ ", 
                             paste(attr(terms(baseline.poly.form.cropyear_omit), 
                                        which = "term.labels"), 
                                   collapse = " + ")),
                        data = crops.train.poly.no_infl)
output_diagnostics(baseline.poly.bc.cropyear_omit, vif = 0)
```

```{r}
multi.form.cropyear_omit = as.formula(Production ~ 
                        (poly(Area, degree = 2, raw = 1) +
                        poly(Annual_Rainfall, degree = 2, raw = 1) +
                        poly(Fertilizer, degree = 2, raw = 1) +
                        poly(Pesticide, degree = 2, raw = 1) +
                        I(sqrt(Area)) +
                        I(sqrt(Annual_Rainfall)) +
                        I(sqrt(Fertilizer)) +
                        I(sqrt(Pesticide)) +
                        I(log(Area)) +
                        I(log(Annual_Rainfall)) +
                        I(log(Fertilizer)) +
                        I(log(Pesticide))) *
                        (Crop + Season + Zone))
multi.full.cropyear_omit = lm(multi.form.cropyear_omit, data = crops.train)
output_diagnostics(multi.full.cropyear_omit, vif = FALSE)
```

```{r}

crops.train.multi.no_infl <- crops.train[!idx_infl_pts(multi.full.cropyear_omit),]

multi.full.cropyear_omit = lm(multi.form.cropyear_omit, data = crops.train.multi.no_infl)
output_diagnostics(multi.full.cropyear_omit, vif = FALSE)
```

```{r}
bc <- boxcox(multi.full.cropyear_omit)
(multi.full.bc.cropyear_omit_lambda <- bc$x[which.max(bc$y)])

multi.full.bc.cropyear_omit = lm(paste("bc_transform(Production, multi.full.bc.cropyear_omit_lambda) ~ ",
                             paste(attr(terms(multi.form.cropyear_omit),
                                        which = "term.labels"),
                                   collapse = " + ")),
                        data = crops.train.multi.no_infl)
output_diagnostics(multi.full.bc.cropyear_omit, vif = FALSE)
```

***

```{r echo=FALSE, message=FALSE}
# multi.aic.backward.bc.cropyear_omit <- step(multi.full.bc.cropyear_omit, direction = "backward", trace = 0)
# output_diagnostics(multi.aic.backward.bc.cropyear_omit, vif = 0)
```

```{r echo=FALSE, message=FALSE}
# multi.bic.backward.bc.cropyear_omit <- step(multi.full.bc.cropyear_omit,
#   direction = "backward", trace = 0,
#   k = log(nrow(crops.train.multi.no_infl))
# )
# output_diagnostics(multi.bic.backward.bc.cropyear_omit, vif = 0)
```

```{r echo=FALSE, message=FALSE}
# multi.aic.forward.bc.cropyear_omit <- step(lm(bc_transform(Production, multi.full.bc.cropyear_omit_lambda) ~ 1, data = crops.train.multi.no_infl),
#                                            scope = as.formula(
#                                              paste("bc_transform(Production, multi.full.bc.cropyear_omit_lambda) ~ ",
#                              paste(attr(terms(multi.form.cropyear_omit),
#                                         which = "term.labels"),
#                                    collapse = " + "))
#                                            ),
#                              direction = "forward", trace = 0)
# output_diagnostics(multi.aic.forward.bc.cropyear_omit, vif = 0)
```

```{r}
multi.bic.forward.bc.cropyear_omit <- step(lm(bc_transform(Production, multi.full.bc.cropyear_omit_lambda) ~ 1, data = crops.train.multi.no_infl),
                                           scope = as.formula(
                                             paste("bc_transform(Production, multi.full.bc.cropyear_omit_lambda) ~ ",
                             paste(attr(terms(multi.form.cropyear_omit),
                                        which = "term.labels"),
                                   collapse = " + "))
                                           ),
                             direction = "forward", trace = 0,
  k = log(nrow(crops.train.multi.no_infl))
)
output_diagnostics(multi.bic.forward.bc.cropyear_omit, vif = 0)
```

## Results

In the following section, we summarize results for the following models explored above:

```{r}
models.cropyear_cont = list(
  baseline.poly.bc.cropyear_cont,
  multi.full.bc.cropyear_cont,
  multi.bic.forward.bc.cropyear_cont
)

models.cropyear_cat = list(
  baseline.poly.bc.cropyear_cat,
  multi.full.bc.cropyear_cat,
  multi.bic.forward.bc.cropyear_cat
)

models.cropyear_omit = list(
  baseline.poly.bc.cropyear_omit,
  multi.full.bc.cropyear_omit,
  multi.bic.forward.bc.cropyear_omit
)

models.cropyear_cont.lambdas = rep(multi.full.bc.cropyear_cont_lambda, 3)
models.cropyear_cont.lambdas[1] = baseline.poly.bc.cropyear_cont_lambda

models.cropyear_cat.lambdas = rep(multi.full.bc.cropyear_cat_lambda, 3)
models.cropyear_cat.lambdas[1] = baseline.poly.bc.cropyear_cat_lambda

models.cropyear_omit.lambdas = rep(multi.full.bc.cropyear_omit_lambda, 3)
models.cropyear_omit.lambdas[1] = baseline.poly.bc.cropyear_omit_lambda

modelNames.cropyear_cont = list(
  "baseline.poly.bc.cropyear_cont",
  "multi.full.bc.cropyear_cont",
  "multi.bic.forward.bc.cropyear_cont"
)

modelNames.cropyear_cat = list(
  "baseline.poly.bc.cropyear_cat",
  "multi.full.bc.cropyear_cat",
  "multi.bic.forward.bc.cropyear_cat"
)

modelNames.cropyear_omit = list(
  "baseline.poly.bc.cropyear_omit",
  "multi.full.bc.cropyear_omit",
  "multi.bic.forward.bc.cropyear_omit"
)
```

```{r}
loocv_rmses.cropyear_cont = rep(0,length(models.cropyear_cont))
train_rmses.cropyear_cont = rep(0,length(models.cropyear_cont))
test_rmses.cropyear_cont = rep(0,length(models.cropyear_cont))
train_adjr2.cropyear_cont = rep(0,length(models.cropyear_cont))
num_predictors.cropyear_cont = rep(0,length(models.cropyear_omit))
for (i in 1:length(models.cropyear_cont)) {
  num_predictors.cropyear_cont[i] = length(coef(models.cropyear_cont[[i]]))
  loocv_rmses.cropyear_cont[i] = get_loocv_rmse(models.cropyear_cont[[i]])
  train_rmses.cropyear_cont[i] = sqrt(mean(resid(models.cropyear_cont[[i]])^2))

  lambda = models.cropyear_cont.lambdas[i]
  
  if(lambda == -1){
    actuals = crops.test.cropyear_cont$Production
  } else {
    actuals = bc_transform(crops.test.cropyear_cont$Production, lambda)
  }

  test_rmses.cropyear_cont[i] = sqrt(
    mean(
      (actuals - predict(models.cropyear_cont[[i]], newdata = crops.test.cropyear_cont))^2
    )
  )
  train_adjr2.cropyear_cont[i] = summary(models.cropyear_cont[[i]])$adj.r.squared
}

loocv_rmses.cropyear_cat = rep(0,length(models.cropyear_cat))
train_rmses.cropyear_cat = rep(0,length(models.cropyear_cat))
test_rmses.cropyear_cat = rep(0,length(models.cropyear_cat))
train_adjr2.cropyear_cat = rep(0,length(models.cropyear_cat))
num_predictors.cropyear_cat = rep(0,length(models.cropyear_omit))
for (i in 1:length(models.cropyear_cat)) {
  num_predictors.cropyear_cat[i] = length(coef(models.cropyear_cat[[i]]))
  loocv_rmses.cropyear_cat[i] = get_loocv_rmse(models.cropyear_cat[[i]])
  train_rmses.cropyear_cat[i] = sqrt(mean(resid(models.cropyear_cat[[i]])^2))
  
  lambda = models.cropyear_cat.lambdas[i]

  if(lambda == -1){
    actuals = crops.test.cropyear_cat$Production
  } else {
    actuals = bc_transform(crops.test.cropyear_cat$Production, lambda)
  }
  
  test_rmses.cropyear_cat[i] = sqrt(
    mean(
      (actuals - predict(models.cropyear_cat[[i]], newdata = crops.test.cropyear_cat))^2
    )
  )
  train_adjr2.cropyear_cat[i] = summary(models.cropyear_cat[[i]])$adj.r.squared
}

loocv_rmses.cropyear_omit = rep(0,length(models.cropyear_omit))
train_rmses.cropyear_omit = rep(0,length(models.cropyear_omit))
test_rmses.cropyear_omit = rep(0,length(models.cropyear_omit))
train_adjr2.cropyear_omit = rep(0,length(models.cropyear_omit))
num_predictors.cropyear_omit = rep(0,length(models.cropyear_omit))
for (i in 1:length(models.cropyear_omit)) {
  num_predictors.cropyear_omit[i] = length(coef(models.cropyear_omit[[i]]))
  loocv_rmses.cropyear_omit[i] = get_loocv_rmse(models.cropyear_omit[[i]])
  train_rmses.cropyear_omit[i] = sqrt(mean(resid(models.cropyear_omit[[i]])^2))
  
  lambda = models.cropyear_omit.lambdas[i]
  
  if(lambda == -1){
    actuals = crops.test.cropyear_omit$Production
  } else {
    actuals = bc_transform(crops.test.cropyear_omit$Production, lambda)
  }
  
  test_rmses.cropyear_omit[i] = sqrt(
    mean(
      (actuals - predict(models.cropyear_omit[[i]], newdata = crops.test.cropyear_omit))^2
    )
  )
  train_adjr2.cropyear_omit[i] = summary(models.cropyear_omit[[i]])$adj.r.squared
}

kable(data.frame("Model" = do.call(c, c(modelNames.cropyear_cont, modelNames.cropyear_cat, modelNames.cropyear_omit)), 
                 "Num Predictors" = c(num_predictors.cropyear_cont, num_predictors.cropyear_cat, num_predictors.cropyear_omit),
                 "LOOCV-RMSE" = c(loocv_rmses.cropyear_cont, loocv_rmses.cropyear_cat, loocv_rmses.cropyear_omit),
                 "Train RMSE" = c(train_rmses.cropyear_cont, train_rmses.cropyear_cat, train_rmses.cropyear_omit), 
                 "Test RMSE" = c(test_rmses.cropyear_cont, test_rmses.cropyear_cat, test_rmses.cropyear_omit),
                 "Train Adj R-Squared" = c(train_adjr2.cropyear_cont, train_adjr2.cropyear_cat, train_adjr2.cropyear_omit)))
```

What these results lead us to conclude is...

```{r}
# TODO: write conclusion
```
